POPLMark goes Concurrent!\\Challenge problems for process calculi and behavioural types

### 2023-10-14 challenges/linearity.tex
minor fixes to subject reduction proof. Preservation of cong still in need of attention

\begin{lemma}[Weakening]\mbox{}
  \label{lemma:weak}
  \begin{enumerate}
  \item If \( \types{\Gamma; \Delta}{P} \) then
    \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
      \item If \( \types{\Gamma; \Delta}{P} \) and \( \tend{\Delta} \) then
    \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
  \end{enumerate}
\end{lemma}
\begin{proof}
  By induction on the given derivations.
\end{proof}
\begin{lemma}[Strengthening]\mbox{}
  \label{lemma:strenD}
  If \( \types{\Gamma; (\Delta, x : T)}{P} \) and
  $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
\end{lemma}
\begin{proof}
  By induction on the last applied rule of the proof of
  $\types{\Gamma; (\Delta, x : T)}{P}$.
  % , for each case inverting $\types{\Gamma; (\Delta, x : T)}{P}$ to
  % the uniquely applicable rule.
  \begin{description}
  \item[\TirName{T-End}] Since $\tend{\Delta}$, we can just reapply
    the rule without $x:T$.
    % [$P = \Pend$] we Apply \TirName{T-Inact}.

@@

  \end{description}
\end{proof}
\begin{lemma}[Preservation for $\equiv$]
  \label{le:presequiv}
  If \( \types{\Gamma;\Delta}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
\end{lemma}
\begin{proof}
  By induction on the structure of a process context involved in the congruence. In the base case where the context is a hole, proceed by case analysis on the \TirName{Sc} rule applied:
  \begin{description}
  \item[Par-Comm/Assoc:] by rearranging sub-derivations noting that
    order does not matter for linear contexts
  \item[Par-Inact] right-to-left by
    strengthening~\ref{lemma:strenD}. Vice-versa, by picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
  \item[Res-Par] Follows Vasco's proof by case analysis on $x : T$ being linear and applying weak and stren accordingly.
  \item[Res-Inact:] use weakening~\ref{lemma:weak} part 2
  \item[Res] order does not matter
  \end{description}
  For the step case, TBD
\end{proof}

>>>>

\begin{lemma}[Weakening]\mbox{}
  \label{lemma:weak}
  \begin{enumerate}
  \item If \( \types{\Gamma; \Delta}{P} \) then
    \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
  \item If \( \types{\Gamma; \Delta}{P} \) % and \( \tend{\Delta} \)
    then
    \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
  \end{enumerate}
\end{lemma}
\begin{proof}
  By induction on the given derivations.
\end{proof}
\begin{lemma}[Strengthening]\mbox{}
  \label{lemma:strenD}
  If \( \types{\Gamma; (\Delta, x : T)}{P} \) and
  $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
\end{lemma}
\begin{proof}
  By induction on the derivation of 
%%  last applied rule of the proof of --AM changed for uniformity
  $\types{\Gamma; (\Delta, x : T)}{P}$.
  % , for each case inverting $\types{\Gamma; (\Delta, x : T)}{P}$ to
  % the uniquely applicable rule.
  \begin{description}
  \item[\TirName{T-End}] Since $\tend{\Delta}$, we can just reapply
    the rule without $x:T$.
    % [$P = \Pend$] we Apply \TirName{T-Inact}.

@@

  \end{description}
\end{proof}
\begin{lemma}[Preservation for $\equiv$]
  \label{le:presequiv}
  If \( \types{\Gamma;\Delta}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
\end{lemma}
\begin{proof}
  By induction on the structure of a process context involved in the congruence. In the base case where the context is a hole, proceed by case analysis on the \TirName{Sc} rule applied:
  \begin{description}
  \item[\TirName {Par-Comm/Assoc}] By rearranging sub-derivations noting that
    order does not matter for linear contexts
  \item[\TirName{Par-Inact}]
    % right-to-left by strengthening~\ref{lemma:strenD}. Vice-versa,
    By picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
  \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear and applying weak and strenghening accordingly.
  \item[\TirName{Res-Inact:}] use weakening~\ref{lemma:weak} part 2
  \item[\TirName{Res}] Noting that order does not matter
  \end{description}
  For the step case, TBD
\end{proof}
